On the Limits of Making Programming Easy

### 2024-12-02 paper.tex
proto-theories etc

\says[\J]{I do not have a perfect answer to this question, but you can think of the basic substrate as a sort of ``user interface physics'' within which specific visual notations for particular problems could be defined.\footnote{Constraint-based systems for specifying user interfaces, such as Cassowary \cite{borning-1997-constraints}, can be seen as related to this idea, although they focus more on rendering the layout for the end-user than on developing a foundational underlying substrate. Additionally, individual constraints are too low-level for a human-friendly UI substrate, although they could well be what such a substrate ``compiles'' to in order to run. This was the subject of much research in the 1990s \cite{myers-1992-uilangs}.} Such a substrate could even be the result of taking a 2D physics engine, stripping it of concepts like mass and rotations, and optimising it for this simpler domain. If we had that, then I think we would make a step towards the ideal substrate for programming. And the substrate would need to live within an actual running materialized system, rather than a static blueprint like a program in source code today.}

@@

@@

>>>>

\says[\J]{I do not have a perfect answer to this question, but you can think of the basic substrate as a sort of ``user interface physics'' within which specific visual notations for particular problems could be defined.\footnote{Constraint-based systems for specifying user interfaces, such as Cassowary \cite{borning-1997-constraints}, can be seen as related to this idea, although they focus more on rendering the layout for the end-user than on developing a foundational underlying substrate. Additionally, individual constraints are too low-level for a human-friendly UI substrate, although they could well be what such a substrate ``compiles'' to in order to run. This was the subject of much research in the 1990s \cite{myers-1992-uilangs}.} Such a substrate could even be the result of taking a 2D physics engine \cite{gabriel-nodate-virtual}, stripping it of concepts like mass and rotations, and optimising it for this simpler domain. If we had that, then I think we would make a step towards the ideal substrate for programming. And the substrate would need to live within an actual running materialized system, rather than a static blueprint like a program in source code today.}

@@

\says[\T]{Well, I would have loved to pursue proto-theories during my PhD---but I fear I would not have been able to defend it!}

\says[\J]{Exactly---there needs to be a way for researchers to get credit for doing this sort of work; until then, it's disincentivised. The \emph{Onward!} tracks at SPLASH provide this opportunity, but there ought to be more venues for it.}

@@

\says[\T]{So your thesis not only proposes a proto-theory, but it also showcases an application of it. Is that something that you'll be able to defend?}

\says[\J]{Let's hope so! I do feel prepared to defend it; we'll find out for sure in one week's time!\footnote{It turned out such a thing did work out after all.}}
